\section*{Local Search}

Local Search is a fundamental optimization framework
\section*{Core Concept}

They are iterative algorithms that each step consider a single current state, and try to 
improve it by moving to one of its neighbor state that has a better evaluation 
score

\begin{itemize}
    \item Search for an improvement is done in local neighborhood only 
    \item Hill climbing, gradient are specific instances of LS
\end{itemize}

Local Search  methods work with complete problem
states/solutions, $s$, i.e., all necessary variables are assigned

\section*{Neighborhoods in discrete domains}

\begin{itemize}
    \item 1-flip N for 0-1 vectors
    \item 2-swap N for permutation vectors 
    \item k-exchange N, $N(s)$ of a state $s$ is the set of states $s'$ that differ
    from $s$ up to $k$ solution components/values
\end{itemize}

A small neighborhood is fast to search at each step, but the search is less powerful.

A large neighborhood is slower to search but can find better solutions.

\section*{Local Search for TSP: Neighborhood Structures}

Local Search for the TSP relies on defining a neighborhood

\subsection*{2-opt}
\begin{tightitemize}
    \item Select two non-adjacent edges, e.g., $(A,B)$ and $(C,D)$.
    \item Remove them, creating two separate paths.
    \item Reconnect the paths as $(A,C)$ and $(B,D)$.
\end{tightitemize}
\textbf{Effect:} Reverses the segment between the two edges, effectively ``uncrossing'' the tour.  

\textbf{Complexity:} $O(N^2)$ possible edge pairs.

\subsection*{3-opt}
\begin{tightitemize}
    \item Delete three edges, splitting the tour into three paths.
    \item Reconnect them in one of $2^3 = 8$ possible configurations (including the original).
\end{tightitemize}
\textbf{Effect:} Allows both reversed and non-reversed segments, enabling solutions unreachable by 2-opt.  

\textbf{Complexity:} $O(N^3)$ possible triples of edges.  

\textbf{Note:} Necessary for Asymmetric TSP (ATSP) where $d(A,B) \neq d(B,A)$.

\subsection*{4-opt (Double Bridge)}
\textbf{Features:}
\begin{itemize}
    \item Does not revert tour segments (useful for ATSP).
    \item Achieves $O(N^2)$ complexity with clever implementation.
    \item Commonly combined with 2-opt for improved performance.
\end{itemize}

\section*{Solution Modification vs Construction}

\begin{itemize}
    \item Solution Modification (Local Search):
    You start with a complete solution (e.g., a full TSP tour) and make local 
    modifications  to iteratively improve it.
    \item Solution Construction:  You start with an empty solution and 
    incrementally build it, one piece at a time, until it is complete.
\end{itemize}
